---
title: "Econometrics Assignment 2"
author: "Ruoheng Du"
date: "04/12/2023"
output: pdf_document
fontsize: 10pt
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plm)
library(dplyr)
```

```{r}
df <- read.csv(file = 'assignment2.csv')
```


### Question 1.1


```{r}
ggplot(data = df, mapping = aes(x = density, y = vio,color = state)) +
  geom_point(stat = 'identity', fill = "#4E84C4") + 
  ylab('Violent Crime Rate') +
  xlab('Density') +
  ggtitle('Density versus Violent Crime Rate') +
  theme_bw() + 
  theme(legend.position = 'none')
```

**DISCUSSION** 
It is hard to discuss the possible result because there's an outlier state, which make it difficult to observe the correlation between density and violent crime rate. Besides, all the data points are clustered in the lower left side, making it hard to observe the result.

```{r}
df[which.max(df$density),]$state
```

**DISCUSSION** 
The outlier is the District of Columbia.

```{r}
ggplot(data = df, mapping = aes(x = log(density), y = log(vio), color = state)) +
  geom_point(stat = 'identity', fill = "#4E84C4")+ 
  ylab('Logged Violent Crime Rate') +
  xlab('Logged Density') +
  ggtitle('Logged Density vs Logged Violent Crime Rate') +
  theme_bw()+ 
  theme(legend.position = 'none')
```

**DISCUSSION** 
It is useful in transforming the variables as now the scatterplot is more accessible. There is no longer an outlier standing way away from the majority of the points, and it can be inferred that there is a positive relationship between density and violent crime rate. 


### Question 1.2


```{r}
ggplot(data = df, mapping = aes(x = shall, y = vio, color = state)) +
  geom_point(stat = 'identity',fill = "#4E84C4") + 
  ylab('Violent Crime Rate') +
  xlab('Shall') +
  ggtitle('Shall versus Violent Crime Rate') +
  theme_bw()+ 
  theme(legend.position = 'none')
```

**DISCUSSION** 
It is possible that there is a negative relationship between shall and violent crime rate, but due to the influence of the outlier discussed above, it is hard to find a clear correlation.


### Question 1.3


```{r}
violent.pd <- pdata.frame(df, index = c("state","year"), 
                          drop.index = TRUE, row.names = TRUE)
head(violent.pd, 3)
```

```{r}
violent.pd <- mutate(violent.pd, lvio = log(vio))
violent.pool <- plm(lvio ~ shall, data = violent.pd, model = 'pooling')
summary(violent.pool, vcov = vcovHC)
```

**DISCUSSION** 
On average, to increase shall by 0.1 of the year, violent crime rate is estimated to decrease by 3.149%. This support the previous view of negative relationship that more shall and less violent crime. With the t-statistic of -2.9338 and p-value of 0.003387, the shall coefficient is statistically significant at any reasonable significance level.


### Question 1.4


```{r}
violent.fe <- plm(lvio ~ shall, data = violent.pd, model = 'within', effect = 
                    'individual')
summary(violent.fe, vcov = vcovHC)
```


**DISCUSSION** 
When adding fixed effects for the different states, on average, a 0.1 increase in shall in the year is associated with a decrease of violent crime rate by 0.33673%. This number is much higher (i.e., smaller effect in absolute sense) compared with the situation that we do not have state fixed effects. 
However, with the t-statistic of -0.8666 and p-value of 0.3863, the shall coefficient is not statistically significant at any reasonable significance level. Besides, the adjusted r-squared value is negative and the p-value of the F-statistic is above 0.1 (i.e., 10% significance level). So, this model is not good.


### Question 1.5


```{r}
violent.fte <- plm(lvio ~ shall, data = violent.pd, model = 'within', effect = 'twoways')
summary(violent.fte, vcov = vcovHC)
```


**DISCUSSION** 
When we control the different characteristics across states and time, on average, a 0.1 increase in shall is associated with a increase of violent crime rate by 0.28157%. This positive sign is different from the previous two signs, indicating that there could be a positive correlation between shall and violent crime rate. However, with the t-statistic of 0.555 and p-value of 0.579, the shall coefficient is not statistically significant at any reasonable significance level. Besides, the adjusted r-squared value is negative and the p-value of the F-statistic is above 0.1 (i.e., 10% significance level). So, this model is not good.


### Question 1.6


```{r}
violent.fe <- plm(lvio ~ shall + rpcpi + rpcui + rpcim + density + pbm1019 + pbm2029 + pwm1019 + pwm2029, data = violent.pd, model = 'within', effect = 'individual')
summary(violent.fe, vcov = vcovHC)
```

```{r}
violent.fte <- plm(lvio ~ shall + rpcpi + rpcui + rpcim + density + pbm1019 + pbm2029 + pwm1019 + pwm2029, data = violent.pd, model = 'within', effect = 'twoways')
summary(violent.fte, vcov = vcovHC)
```


**DISCUSSION** 
Beta1 becomes negative again and it changes a lot compared with the original regression, but it is not statistically significant at any reasonable significance level. 
Compared with the pooled model, the four regression models with state/time fixed effects are more credible because they control for omitted variable bias. Compared with FE and TFE without controls (i.e., in Question 1.4 & 1.5), which have negative adjusted r-squared values, these with controls (i.e., in Question 1.6) have higher adjusted r-squared values of about 17% and 9%. However, the t-statistics of the shall coefficient in these four models with fixed effects indicate that they are all statistically insignificant. Overall, I would say that there is no significant effect of shall-issue laws on the violent crime rate.
